---
title: "EDA"
author: "Pei-Hsin Lin"
date: "3/30/2021"
output: pdf_document
---

```{r}
library(tidyverse)
```


# Load & clean data
```{r}
data <- read_csv("data/Project_Data_final.csv")
data <- na.omit(data) # remove data row w/ NA
inf <- data$X1[is.infinite(data$pop_density_sqmi)]
data <- data[-inf,]
nrow(data)
names(data)

data <- data[,c(3,4,6,9,19,22,23,24,25,27)] # remove redundant variables
# unique(data$Region) 
unique(data$Division) 
head(data)
```


# EDA

```{r}
# data$Region <- as.numeric(as.factor(data$Region))
data$Division_num <- as.numeric(as.factor(data$Division))
# Regions:
# 1 = Midwest
# 2 = Northeast
# 3 = South
# 4 = West
# 
# Divisions:
# 1 = East North Central
# 2 = East South Central
# 3 = Middle Atlantic
# 4 = Mountain
# 5 = New England
# 6 = Pacific
# 7 = South Atlantic
# 8 = West North Central
# 9 = West South Central

X<-data[,-c(3,10)] # remove date (non-numeric)
names(X)
M<-cor(X)
library(corrplot)
corrplot(M, method="ellipse", tl.cex = 0.3)
corrplot(M, method="number", tl.cex=.8,number.cex=.8,tl.srt=30)
```
```{r}
hist(data$median_listing_price)
par(mfrow=c(2,2))
plot(data$median_listing_price~.,data=X)
```


```{r}
par(mfcol=c(4,2))
hist(data$median_listing_price,col="red",main="median_listing_price",xlab = NULL)
hist(data$total_listing_count,col="green",main="total_listing_count",xlab = NULL)
hist(data$POP_ESTIMATE_2019,col="blue",main="POP_ESTIMATE_2019",xlab = NULL)
hist(data$cases_per_1000,col="pink",main="cases_per_1000",xlab = NULL)
hist(data$deaths_per_1000,col = 'purple',main = 'deaths_per_1000',xlab = NULL)
hist(data$county_land_area,col='grey',main='county_land_area',xlab = NULL)
hist(data$pop_density_sqmi,col='brown',main='county_land_area',xlab = NULL)
hist(data$Median_Household_Income_2019,col='orange',main='Median_Household_Income_2019',xlab = NULL)
```


```{r message=F}
ggpairs(data,lower = list(continuous = wrap("points", alpha = 0.05)))
```

## log transform
```{r}
names(data)
data[,-c(3,10)] %>% 
  mutate_at(1:8, list(log = ~ log(.))) ->
  data_log
data_log <- as.data.frame(Map(function(x) replace(x, is.infinite(x), 0), data_log))
data_log <- data_log[,c(9:16)]# only include the logged variables
data_log <- cbind('date'=data$date,'Division'=data$Division,data_log)
# data_log <- cbind('Division'=data$Division,data_log)
data_log <- na.omit(data_log)
names(data_log)
data_log$Division <- as.factor(data_log$Division)
```

## pairplot
```{r message=F}
library(GGally)
ggpairs(data_log,lower = list(continuous = wrap("points", alpha = 0.05))) # some still skewed..
```

```{r}
par(mfcol=c(4,2))
hist(data_log$median_listing_price,col="red",main="median_listing_price",xlab = NULL)
hist(data_log$total_listing_count,col="green",main="total_listing_count",xlab = NULL)
hist(data_log$POP_ESTIMATE_2019,col="blue",main="POP_ESTIMATE_2019",xlab = NULL)
hist(data_log$cases_per_1000,col="pink",main="cases_per_1000",xlab = NULL)
hist(data_log$deaths_per_1000,col = 'purple',main = 'deaths_per_1000',xlab = NULL)
hist(data_log$county_land_area,col='grey',main='county_land_area',xlab = NULL)
hist(data_log$pop_density_sqmi,col='brown',main='county_land_area',xlab = NULL)
hist(data_log$Median_Household_Income_2019,col='orange',main='Median_Household_Income_2019',xlab = NULL)
```


# Model
## COVID effect on listing price
```{r}
lm.covid.fit<-lm(median_listing_price_log~cases_per_1000_log,data=data_log)
summary(lm.covid.fit)
lm.covid.fit2<-lm(median_listing_price_log~deaths_per_1000_log,data=data_log)
summary(lm.covid.fit2)

lm.covid.fit3<-multinom(median_listing_price_clas~cases_per_1000_log,data=data_log)
summary(lm.covid.fit3)
lm.covid.fit4<-multinom(median_listing_price_clas~deaths_per_1000_log,data=data_log)
summary(lm.covid.fit4)
```

```{r}
library(lmtest)

par(mfcol=c(2,2))
bptest(lm.covid.fit)
plot(lm.covid.fit, which = 1)

bptest(lm.covid.fit2)
plot(lm.covid.fit2, which = 1)

bptest(lm.covid.fit3)
plot(lm.covid.fit3, which = 1)

bptest(lm.covid.fit4)
plot(lm.covid.fit4, which = 1)
```


## Regression
```{r}
lm.fit<-lm(median_listing_price_log~.,data=data_log)
summary(lm.fit) # pop_density_sqmi_log NA due to strong collinearity # high R2
```


## logistic regression
```{r}
data_log$median_listing_price_clas <- data$median_listing_price
mlp_mean <- mean(data_log$median_listing_price_clas)
mlp_std <- sd(data_log$median_listing_price_clas)
ifelse(data_log$median_listing_price_clas > mlp_mean+mlp_std, "High",
       ifelse(data_log$median_listing_price_clas < mlp_mean-mlp_std,"Low","Median")) ->
  data_log$median_listing_price_clas
data_log$median_listing_price_clas <- as.numeric(as.factor(data_log$median_listing_price_clas))
head(data_log)

library(nnet)
multi_log.fit <- multinom(median_listing_price_clas~.-median_listing_price_log,data=data_log)
summary(multi_log.fit)
```


## Ridge/LASSO
```{r}
library(glmnet)
set.seed(1)
names(train_data)
x <- model.matrix(median_listing_price_log~.-median_listing_price_clas,data = data_log)[,-1]
y <- data_log$median_listing_price_log 
lasso <- glmnet(x, y, alpha=1) 
set.seed(1)
lasso.cv.10Fold <- cv.glmnet(x,y,alpha=1) 
lasso.best.lambda <- lasso.cv.10Fold$lambda.min 
lasso.min.mse <- min(lasso.cv.10Fold$cvm)
cbind("Best Lambda"=lasso.best.lambda, 
      "Log(Lambda)"=log(lasso.best.lambda), 
      "Best 10FCV MSE" = lasso.min.mse)
lasso.coef <- coef(lasso, s=lasso.best.lambda) 
lasso.coef
```

remove case and pop_dens
final predictors: date, Division, POP_ESTIMATE_2019_log, Median_Household_Income_2019_log, total_listing_count_log, deaths_per_1000_log, county_land_area_log

## final reg
```{r}
reg.final <- lm(median_listing_price_log ~ date + Division + POP_ESTIMATE_2019_log + 
                  Median_Household_Income_2019_log + total_listing_count_log + 
                  deaths_per_1000_log + county_land_area_log, data=data_log)
summary(reg.final)
```

```{r}
library(nnet)
multi_log.final <- multinom(median_listing_price_clas ~ date + Division + POP_ESTIMATE_2019_log + 
                              Median_Household_Income_2019_log + total_listing_count_log + 
                              deaths_per_1000_log + county_land_area_log, data=data_log)
summary(multi_log.final)
```


## 80/20 split
```{r}
set.seed(123)
train <- sample(nrow(data_log),0.8*nrow(data_log))
train_data <- data_log[train,-c(7,10)]
test_data <- data_log[-train,-c(7,10)]
nrow(train_data)
nrow(test_data)
names(train_data)
```
## train/test regression
```{r}
reg.train <- lm(median_listing_price_log ~ date + Division + POP_ESTIMATE_2019_log +
                  Median_Household_Income_2019_log + total_listing_count_log + 
                  deaths_per_1000_log + county_land_area_log, data=train_data)
```

### Accuracy of Train data
```{r}
pred_train <- predict(reg.final,train_data)
RMSE_train <- sqrt(sum((pred_train-train_data$median_listing_price_log)^2)/nrow(train_data))
RMSE_train
```

### Accuracy of Test data  
```{r}
pred_test <- predict(reg.final,test_data)
RMSE_test <- sqrt(sum((pred_test-test_data$median_listing_price_log)^2)/nrow(test_data))
RMSE_test
```

## train/test logistic
### Table of Coefficient,standard error, z statand, p value for Probability
```{r}
multi_log.train <- multinom(median_listing_price_clas ~ date + Division + POP_ESTIMATE_2019_log + 
                              Median_Household_Income_2019_log + total_listing_count_log + 
                              deaths_per_1000_log + county_land_area_log, data=train_data)
muti_log.output <- summary(multi_log.train)
z <- muti_log.output$coefficients/muti_log.output$standard.errors
p <- (1-pnorm(abs(z),0,1))*2
Pquality <- rbind(muti_log.output$coefficients[2, ],muti_log.output$standard.errors[2, ],z[2, ],p[2, ])
rownames(Pquality) <- c("Coefficient","Std. Errors","z stat","p-value")
knitr::kable(Pquality)
```

### Accuracy of Train data
```{r}
pred <- predict(multi_log.final,train_data,na.action=na.pass, type="probs")
pred_class <- predict(multi_log.final,train_data)
caret::confusionMatrix(as.factor(pred_class),as.factor(train_data$median_listing_price_clas))
```

### Accuracy on test data  
```{r}
pred_test<- predict(multi_log.final,test_data,na.action=na.pass, type="probs")
pred_test_class <- predict(multi_log.final,test_data)
caret::confusionMatrix(as.factor(pred_test_class),as.factor(test_data$median_listing_price_clas))
```






