---
title: "EDA"
author: "Pei-Hsin Lin"
date: "3/30/2021"
output: pdf_document
---

```{r}
library(tidyverse)
data <- read_csv("data/project_data.csv")[,-1]
data <- na.omit(data)
head(data)
names(data)
unique(data$date)
unique(data$county_fips) # 986
unique(data$state) # 51
```


```{r}
data$region <- 0
for (i in 1:nrow(data)) {
  if (data$state[i] %in% c("connecticut","maine","massachusetts","new hampshire","rhode island", "vermont")) {
    data$region[i] <- 1
  } else if (data$state[i] %in% c("new jersey", "new york", "puerto rico")) {
    data$region[i] <- 2
  } else if (data$state[i] %in% c("delaware", "district of columbia", "maryland", "pennsylvania", "virginia", "west virginia")) {
    data$region[i] <- 3
  } else if (data$state[i] %in% c("alabama", "florida", "georgia", "kentucky", "mississippi", "north carolina", "south carolina", "tennessee")) {
    data$region[i] <- 4
  } else if (data$state[i] %in% c("illinois", "indiana", "michigan", "minnesota", "ohio", "wisconsin")) {
    data$region[i] <- 5
  } else if (data$state[i] %in% c("arkansas", "louisiana", "new mexico", "oklahoma", "texas")) {
    data$region[i] <- 6
  } else if (data$state[i] %in% c("iowa", "kansas", "missouri", "nebraska")) {
    data$region[i] <- 7
  } else if (data$state[i] %in% c("colorado", "montana", "north dakota", "south dakota", "utah", "wyoming")) {
    data$region[i] <- 8
  } else if (data$state[i] %in% c("arizona", "california", "hawaii", "nevada", "american samoa", "guam", "northern mariana islands")) {
    data$region[i] <- 9
  } else if (data$state[i] %in% c( "alaska", "idaho", "oregon", "washington")) {
    data$region[i] <- 10
  }
}
data$region <- as.factor(data$region)
# Region I: Connecticut, Maine, Massachusetts, New Hampshire, Rhode Island, Vermont
# Region II: New Jersey, New York, Puerto Rico, US Virgin Islands
# Region III: Delaware, District of Columbia, Maryland, Pennsylvania, Virginia, West Virginia
# Region IV: Alabama, Florida, Georgia, Kentucky, Mississippi, North Carolina, South Carolina, Tennessee
# Region V: Illinois, Indiana, Michigan, Minnesota, Ohio, Wisconsin
# Region VI: Arkansas, Louisiana, New Mexico, Oklahoma, Texas
# Region VII: Iowa, Kansas, Missouri, Nebraska
# Region VIII: Colorado, Montana, North Dakota, South Dakota, Utah, Wyoming
# Region IX: Arizona, California, Hawaii, Nevada, American Samoa, Guam, Northern Mariana Islands
# Region X: Alaska, Idaho, Oregon, Washington
```



# EDA
1. Does the monthly amount of covid cases affect the rental price over states?
- Regression: rental~covid_cases

```{r}
library(ggplot2)
```

```{r}
# ggplot(aes(x=date, y=cases), data=data) + 
#   geom_point() + geom_smooth()

ggplot(aes(x=date, y=cases), data=data) + 
  geom_point() + geom_smooth() +
  facet_grid(cols = vars(region)) +
  theme(axis.text.x=element_text(angle=45))
```

2. When does the trend of covid start to slow down? 
   Does monthly housing inventory decrease when the covid cases increase?
- 
- 
```{r}
data%>%
  group_by(date,region)%>%
  summarise(Covid_Cases=mean(cases))%>%
  ungroup()%>%
  ggplot(aes(
    x=date,
    y=Covid_Cases,
    colour=region,
    group=region,
    shape=region
    ))+
  geom_line()+
  geom_point(size=4)
```

```{r}
data%>%
  group_by(date,region)%>%
  summarise(mlp=mean(median_listing_price))%>%
  ungroup()%>%
  ggplot(aes(
    x=date,
    y=mlp,
    colour=region,
    group=region,
    shape=region
    ))+
  geom_line()+
  geom_point(size=4)

data%>%
  group_by(date,region)%>%
  summarise(mlp=mean(active_listing_count))%>%
  ungroup()%>%
  ggplot(aes(
    x=date,
    y=mlp,
    colour=region,
    group=region,
    shape=region
    ))+
  geom_line()+
  geom_point(size=4)

```
```{r}
data%>%
  group_by(date,region)%>%
  summarise(Covid_Cases=mean(cases), mlp=mean(median_listing_price))%>%
  ungroup()%>%
  ggplot(aes(
        x = Covid_Cases,
        y = mlp,
        col = region
        )) +
       geom_point(mapping = aes(size = mlp)) +
      geom_point(aes(size = Covid_Cases)) +
      scale_size(range = c(1, 20), guide = FALSE) +
      scale_x_discrete(aes(label = Covid_Cases)) +
      scale_color_brewer(palette = "Set3")
```

3a. Time series for rental price (Year → Month)
Can the rental price of 2020 be predicted by the former rental price? 
If not, it seems like covid do influence the rate. 
We can take the next step to set up regression analysis for rental price prediction.

```{r}

```
 
3b. Regression (rental price prediction)
-- regression
-- regression tree
Then, we would like to know whether covid significantly affects the rental price yearly?
- yearly_rental_price ~ covid_case(1/0)
If yes, we could also explore the effect of covid on monthly rental price for the whole U.S.
- monthly_rental_price ~ covid_case(quant)
 
Then, we want to know if there are any other factors contributing to the monthly rental price 
for the US and each state.
- monthly_rental_price ~ population(quant or rank)+covid_case(quant)+covid_death
- monthly_rental_price ~ state+population(quant or rank)+covid_case(quant)+covid_death
 
3c. Classification:
We can then divide monthly rental price into two/three levels (high,medium,low), and see 
how the factors affect the rental price level.
-- logistic regression
-- classification tree
- yearly_rental_price_level ~ covid_case(1/0)
- monthly_rental_price_level ~ covid_case(quant)
- monthly_rental_price_level ~ population(quant or rank)+covid_case(quant)+covid_death
- monthly_rental_price_level ~ state+population(quant or rank)+covid_case(quant)+covid_death











